For submission to the Journal of Integrated Pest Management

### 2021-10-04 editorial_reviews/response_editors_ofv_fife_2021_R3.Rmd

- The Spanish abstract addition is not warranted and I am confused why you added this at this late stage? The staff editors have suggested “So, my suggestion would be just to assume the other [English] abstract is ok.”. $\rightarrow$ **the authors added the Spanish abstract in order to be more inclusive of the Latin American researchers who study Orchid fleck virus and similar pathogens in the citrus-growing regions Mexico and South America. We believe that the abstract is in compliance with the directions described in the 'Language' section under 'Manuscript Preparation' for the Journal of Integrated Pest Management: https://academic.oup.com/jipm/pages/Manuscript_Preparation, which indicates that a "...A second abstract in a second language is permitted." The authors did not intend to cause concern by adding the abstract at a later stage of editing, because JIPM's Manuscript Preparation page says "Authors are responsible for the accuracy of non-English abstracts". The authors believe that our abstract will be found to be an accurate translation, faithful to the English version.**

- Line 91: ‘OFV is a bacilliform’, here and throughout, while not a big issue generally it is preferred that a sentence is not started with and acronym. Consider ‘The virus is a bacilliform’ $\rightarrow$ **moved acronyms from beginnings of sentences**

- Line 94: are the only which have been- consider improving the wording $\rightarrow$ **changed to "Only Flat mites (Trombidiformes: Tenuipalpidae) from the genus *Brevipalpus* are known to transmit dichorhaviruses..."**

- Line 96: on the infected plant, remove infected $\rightarrow$ **removed**

@@

- Line 140: define “HTS”, just because I know what that is in English does not mean someone else will know what that is. $\rightarrow$ **changed to "High Throughput Sequencing (HTS)"**

- Line 149: HTS results ‘for samples obtained’ from Leon County revealed that *L. muscari* $\rightarrow$ **reworded**

- Line 199: identical in gene order, content, and genome sequence – I think you mean genome sequence? I am not sure what is meant by gene order and content, it is an identical genome, correct? $\rightarrow$ **removed repeated content**

Please ensure all authors have reviewed the document. $\rightarrow$ **All authors have reviewed the document, some sections which implied that OFV may pose a risk to citrus were removed and/or reworded to avoid speculation, as per the request of the review board of one of the authors.**

### 2021-10-07 ofv_fife_2021.Rmd
complying with review boards and editor comments
added recommended edits

Mite taxonomy is complicated by cryptic species complexes which occur in many plant-feeding groups of the Acari [@Arthur2011; @Skoracka2010; @Skoracka2013; @Umina1999], including tenuipalpid mites from the genus *Brevipalpus* [@Navia2013]. The commonly used phase-contrast microscopy is insufficient to detect some diagnostic characters for separation of cryptic species, instead best practices recommend the combination of Differential Interference Contrast (DIC) Microscopy and Scanning Electron Microscopy along with molecular methods to separate cryptic species [@Beard2015]. The flat mites collected were initially suspected to belong to *B. californicus* after inspection with phase contrast microscopy. Subsequent observation via DIC microscopy at FDACS agreed with this tentative identification. Unfortunately, the *B. californicus* s.l. species group, *sensu* @Baker1987 is suspected to contain cryptic species [@Childers2011; @Rodrigues2013]. New mite samples were collected from symptomatic liriopogons and *A. elatior* in Leon County and sent to USDA-ARS's Electron and Confocal Microscopy Unit for analysis. Three mite species were recovered and examined under cryo-scanning electron microscopy (Cryo-SEM): *B. californicus* s.l. (`r citef("fig_3")`), *B. obovatus* Donnadieu and *B. confusus* Baker. The recent report of OFV in the US is thought to be @Ko1985 which describes nuclear inclusions caused by an undescribed bacilliform rhabdovirus in *Brassia* orchids. The significance of this report is their description of the spoke-wheel configurations of the viral particles [@Ko1985], a sign typically associated with OFV infection [@Chang1976]. Unfortunately, this article made no mention of mites or further investigations of the virus. The first report of OFV in the continental US was @Bratsch2015, who confirmed the presence of OFV in *Phalaenopsis* hybrids using Transmission Electron Microscopy of ultrathin sections of plant tissue as well as molecular sequence analysis. They also discuss the association of OFV with *Brevipalpus* mites, but the authors did not make a conclusive species identification beyond suggesting that the mite vector belonged to the *B. californicus* group, referring to @Kondo2003's publication [@Bratsch2015]. Later reports of OFV described OFV infecting a previously undescribed Nolinoidaea hosts in Australia [@Dietzgen2018a; @Mei2016], including *Liriope spicata* (Thunb.) Lour, a different species of liriopogon than those identified from the Florida sites. We are not aware of any reports of OFV infecting liriopogons, *A. elatior* nor other Nolinoidaea in the US. Although @Peng2013 had mentioned an association between *B. californicus* and *A. elatior*, they never reported symptoms of OFV-Orc in this plant. We believe that our findings indicate the first report of OFV-Orc infecting ornamental Nolinoidaea in Florida, and possibly the US. This publication also marks the first reports of *A. elatior* and *Ophiopogon* spp. as natural hosts of OFV-Orc. There are two orchid strains of OFV (OFV-Orc1 and OFV-Orc2), and two citrus strains (OFV-Cit1 and OFV-Cit2) [@Roy2020; @Beltran-Beltran2020]. The OFV strains detected in Florida are identical in genome sequence to the orchid strains of OFV infecting citrus in Hawaii, Mexico, Colombia, and South Africa [@Roy2020; @Beltran-Beltran2020]. Both OFV-Orc1 and OFV-Orc2 infect citrus [@Roy2020], but none of the citrus  strains have been reported from any orchid species. It is important to note the uncertainty surrounding the vector for OFV-Orc. There are three mite species which have been recovered from OFV-Orc infected plants: *B. obovatus*, and *B. confusus* and *B. californicus* s.l., but only *B. californicus* has been described as a vector of OFV. Even so, these types of questions require future study to determine the potential of nolinoidaea to citrus transmission. Best practices for integrated pest management have not been created for controlling *Brevipalpus* mites on these ornamentals, but methods designed to control *Brevipalpus* in other systems may be applicable. The most common method used to control *Bervipalpus* are synthetic acaricides [@Andrade2010; @Andrade2019]. Unfortunately, some acaricides and their residues can harm beneficial predatory mites as well [@Fernandez2017], even at low doses [@Havasi2021], and mixing different chemistries can be detrimental for mite control [@Vechia2018]. In addition, pesticide resistance has been reported in various *Brevipalpus* populations [@Rocha2021; @Alves2000; @Campos2002; @Omoto2000], due to exposure to pesticides used to control other arthropod pests [@Vechia2021]. In addition, predatory mites [@Argolo2020; @Chen2006], entomopathogenic fungi [@RossiZalaf2008; @Magalhaes2005; @Pena2015; @Revynthi2019] have shown promise for controlling other *Brevipalpus* mites. Moreover, it is often possible to integrate different control techniques for improved management, such as combining predatory mites with compatible acaricides and entomopathogenic fungi [@Reddy2001; @Midthassel2016; @Andrade2019]. In conclusion, detecting OFV in Florida represents a concern for horticulturists who grow orchids, *Liriope*, *Ophiopogon*, or other susceptible Asparagaceae species which are commonly used in landscaping. Florida is also home to a plethora of native and naturalized orchid species, many of which are threatened, including cultivated *Vanilla* in southern Florida [@Chambers2019] and the famous Ghost Orchid, [*Dendrophylax lindenii* (Lindl.) Benth. ex Rolfe]. Citrus leprosis was present in Florida during the 1860’s and almost eradicated by the mid-1960s [@Childers2003; @Knorr1968a; @Knorr1968b]. An examination of herbarium specimens of Florida citrus found that this historical virus, Citrus leprosis dichorhavirus-N0, is distantly related to the modern isolates of OFV [@Hartung2015; @Kitajima2011a; @Roy2020]. The recent detection of OFV-Orc1 in South Africa [@Cook2019] in *C. sinensis* (Navel and Valencia orange) and OFV-Orc2 in Hawaii [@Velarde2021] in *C. reticulata* (mandarin) and *C. jambhiri* (rough lemon) associated with leprosis-like symptoms highlights the potential threat of different isolates of OFV on citrus. *B. californicus*, *B. yothersi*, and *B. obovatus* are all present in Florida [@Childers2003; @Akyazi2017], and are difficult to identify by non-experts, or without advanced methodologies. DNA barcoding [@Armstrong2005] or a similarly simple and accurate method for identification of these mite complexes is vital to identify mite populations which need to be monitored or controlled. By doing so, we can determine the risk OFV-Orc represents for the native plants, agriculture and the ornamental/landscaping industries of Florida and the surrounding regions.

>>>>

Mite taxonomy is complicated by cryptic species complexes which occur in many plant-feeding groups of the Acari [@Arthur2011; @Skoracka2010; @Skoracka2013; @Umina1999], including tenuipalpid mites from the genus *Brevipalpus* [@Navia2013]. The commonly used phase-contrast microscopy is insufficient to detect some diagnostic characters for separation of cryptic species, instead best practices recommend the combination of Differential Interference Contrast (DIC) Microscopy and Scanning Electron Microscopy along with molecular methods to separate cryptic species [@Beard2015]. The flat mites collected were initially suspected to belong to *B. californicus* after inspection with phase contrast microscopy. Subsequent observation via DIC microscopy at FDACS agreed with this tentative identification. Unfortunately, the *B. californicus* s.l. species group, *sensu* @Baker1987 is suspected to contain cryptic species [@Childers2011; @Rodrigues2013]. New mite samples were collected from symptomatic liriopogons and *A. elatior* in Leon County and sent to USDA-ARS's Electron and Confocal Microscopy Unit for analysis. Three mite species were recovered and examined under cryo-scanning electron microscopy (Cryo-SEM): *B. californicus* s.l. (`r citef("fig_3")`), *B. obovatus* Donnadieu and *B. confusus* Baker. The recent report of OFV in the US is thought to be @Ko1985 which describes nuclear inclusions caused by an undescribed bacilliform rhabdovirus in *Brassia* orchids. The significance of this report is their description of the spoke-wheel configurations of the viral particles [@Ko1985], a sign typically associated with OFV infection [@Chang1976]. Unfortunately, this article made no mention of mites or further investigations of the virus. The first report of OFV in the continental US was @Bratsch2015, who confirmed the presence of OFV in *Phalaenopsis* hybrids using Transmission Electron Microscopy of ultrathin sections of plant tissue as well as molecular sequence analysis. They also discuss the association of OFV with *Brevipalpus* mites, but the authors did not make a conclusive species identification beyond suggesting that the mite vector belonged to the *B. californicus* group, referring to @Kondo2003. Later reports of OFV described OFV infecting a previously undescribed Nolinoidaea hosts in Australia [@Dietzgen2018a; @Mei2016], including *Liriope spicata* (Thunb.) Lour, a different species of liriopogon than those identified from the Florida sites. We are not aware of any reports of OFV infecting liriopogons, *A. elatior* nor other Nolinoidaea in the US. Although @Peng2013 had mentioned an association between *B. californicus* and *A. elatior*, they never reported symptoms of OFV-Orc in this plant. We believe that our findings indicate the first report of OFV-Orc infecting ornamental Nolinoidaea in Florida, and possibly the US. This publication also marks the first reports of *A. elatior* and *Ophiopogon* spp. as natural hosts of OFV-Orc. There are two orchid strains of OFV (OFV-Orc1 and OFV-Orc2), and two citrus strains (OFV-Cit1 and OFV-Cit2) [@Roy2020; @Beltran-Beltran2020]. The OFV strains detected in Florida are identical in genome sequence to the orchid strains of OFV infecting citrus in Hawaii, Mexico, Colombia, and South Africa [@Roy2020; @Beltran-Beltran2020]. Both OFV-Orc1 and OFV-Orc2 infect citrus [@Roy2020], but none of the citrus  strains have been reported from any orchid species. It is important to note the uncertainty surrounding the vector for OFV-Orc. There are three mite species which have been recovered from OFV-Orc infected plants: *B. obovatus*, and *B. confusus* and *B. californicus* s.l., but only *B. californicus* has been described as a vector of OFV. Even so, these types of questions require future study to determine the potential of nolinoidaea to citrus transmission. Best practices for integrated pest management have not been created for controlling *Brevipalpus* mites on these ornamentals, but methods designed to control *Brevipalpus* in other systems may be applicable. The most common method used to control *Bervipalpus* are synthetic acaricides, such as spirodiclofen and cyflumetofen [@Vechia2018; @Andrade2010; @Andrade2019]. Unfortunately, some acaricides and their residues can harm beneficial predatory mites as well [@Fernandez2017], even at low doses [@Havasi2021], and mixing different chemistries can be detrimental for mite control [@Vechia2018]. Furthermore, pesticide resistance has been reported in various *Brevipalpus* populations [@Rocha2021; @Alves2000; @Campos2002; @Omoto2000], due to exposure to pesticides used to control other arthropod pests [@Vechia2021]. In addition, phytoseiid predatory mites such as *Amblyseius largoensis* (Muma) and *Galendromus helveolus* (Chant) [@Argolo2020; @Chen2006], entomopathogenic fungi [@RossiZalaf2008; @Magalhaes2005; @Pena2015; @Revynthi2019] have shown promise for controlling other *Brevipalpus* mites. Moreover, it is often possible to integrate different control techniques for improved management, such as combining predatory mites with compatible acaricides and entomopathogenic fungi [@Reddy2001; @Midthassel2016; @Andrade2019]. In conclusion, detecting OFV in Florida represents a concern for horticulturists who grow orchids, *Liriope*, *Ophiopogon*, or other susceptible Asparagaceae species which are commonly used in landscaping. Florida is also home to a plethora of native and naturalized orchid species, many of which are threatened, including cultivated *Vanilla* in southern Florida [@Chambers2019] and the famous Ghost Orchid, [*Dendrophylax lindenii* (Lindl.) Benth. ex Rolfe]. Citrus leprosis was present in Florida during the 1860’s and almost eradicated by the mid-1960s [@Childers2003; @Knorr1968a; @Knorr1968b]. An examination of herbarium specimens of Florida citrus found that one strain of this historical virus, Citrus leprosis dichorhavirus-N0, is distantly related to the modern isolates of OFV [@Hartung2015; @Kitajima2011a; @Roy2020]. The recent detection of OFV-Orc1 in South Africa [@Cook2019] in *C. sinensis* (Navel and Valencia orange) and OFV-Orc2 in Hawaii [@Velarde2021] in *C. reticulata* (mandarin) and *C. jambhiri* (rough lemon) associated with leprosis-like symptoms highlights the potential threat of different isolates of OFV on citrus. *B. californicus*, *B. yothersi*, and *B. obovatus* are all present in Florida [@Childers2003; @Akyazi2017], and are difficult to identify by non-experts, or without advanced methodologies. DNA barcoding [@Armstrong2005] or a similarly simple and accurate method for identification of these mite complexes is vital to identify mite populations which need to be monitored or controlled. By doing so, we can determine the risk OFV-Orc represents for the native plants, agriculture and the ornamental/landscaping industries of Florida and the surrounding regions.
